---
title: "Enrichment Workflow Step Dev"
author: "Joseph"
date: "06/05/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(jsonlite)
library(tidyr)
library(dplyr)
library(EnrichmentBrowser)
library(stringr)
library(R.utils)

```

To do: 

* Add observed list to enrichment table
* Create Inverse Enrichment Tables
* 


After chat with Az:

* make a workflow step
* write some acceptance tests
* get go ahead from Peppe to get BKG development going

# Experiment Parsing:

```{r small experiment}
library(EnrichmentAnalysisStepR)
experiment_home = "/home/joseph/experiments/her_data_gsea_test"
gmt_folder = "/home/joseph/experiments/BKG_gmt_output_human/9606"

results = enrichment_workflow_step(experiment_home, 
                                   gmt_folder = gmt_folder,
                                   output_folder = file.path(experiment_home,"transform"),
                                   by = "Protein")


```

```{r}
library(EnrichmentAnalysisStepR)
experiment_home = "/home/joseph/experiments/her_data_gsea_test"
gmt_folder = "/home/joseph/experiments/BKG_gmt_output_human/9606"



results = enrichment_workflow_step(experiment_home, 
                                   gmt_folder = gmt_folder,
                                   output_folder = file.path(experiment_home,"transform"),
                                   by = "Protein",
                                   method = "padog")
```

```{r}
plot(results$`C - P`$BiologicalProcess$mean.abs.t0,
     -log10(results$`C - P`$BiologicalProcess$pval))
```

```{r large experiment}
experiment_home = "/home/joseph/experiments/mouse_pxd_ida"
gmt_folder = "/home/joseph/experiments/BKG_gmt_output_mouse"
results = enrichment_workflow_step(experiment_home, 
                                      gmt_folder = gmt_folder,
                                      output_folder = file.path(experiment_home,"transform"),
                                      by = "Protein")


```


```{r}
experiment_home = "/home/joseph/experiments/ups_data_test/"
results = enrichment_workflow_step(experiment_home, 
                                      gmt_folder ="./ckg_gmts_by_gene",  
                                      output_folder = file.path(experiment_home,"transform"),
                                      by = "Gene")


```

# Visualisation Experiments
## Scatter

```{r}

#results$`C - P`$Biological_process

library(plotly)

scatter_enrichment_results <- function(enrichment_table){
  
  plotting_df = as.data.frame(enrichment_table)
  plotting_df$log10pval = -1*log10(plotting_df$ADJ.PVAL)
  
  p <- ggplot(plotting_df, aes(x = NES, y = log10pval, label = GENE.SET)) + 
    geom_point(show.legend = FALSE, aes(size = SIZE)) +
    labs(title = "Scatter Enrichment Results", 
         x = "Normalized Enrichment Score",
         y = "-log10(P-Value)") +
    geom_hline(yintercept=-1*log10(0.05), linetype="dashed")+
    theme_minimal()
  
  fig <- ggplotly(p, tooltip = c("label", "x", "y"))
  fig$x$layout$showlegend = FALSE
  fig
  
  
}

scatter_enrichment_results(results$`Kidney - Muscle`$Pathway)
scatter_enrichment_results(results$`Kidney - Muscle`$Cellular_component)
scatter_enrichment_results(results$`Kidney - Muscle`$Publication)
```

## Bar Plot

```{r}
# We're going to need the protein expression results
enr_results = results$`Kidney - Muscle`$Publication
plotting_df = enr_results[enr_results$ADJ.PVAL<0.001,]
plotting_df <- plotting_df[order(plotting_df$NES),]

ggplot(plotting_df, aes(NES, GENE.SET, fill = ADJ.PVAL)) + 
  geom_col() +
  scale_fill_continuous(low="red", 
                        high="blue", 
                        name = "ADJ.PVAL",
                        guide=guide_colorbar(reverse=TRUE)) +
  theme_minimal()


```


# add observed to 

# create inverse enrichment table

```{r}
fix_string_list <- function(string_list){
  string_list <- gsub("\\[","", string_list)
  string_list <- gsub("\\]","", string_list)
  string_list <- gsub("\\'","", string_list)
  string_list <- gsub("\\s","", string_list)
  list <- str_split(string_list, ",")
  unlist(list)
}

get_inv_enrichment_table <- function(enr_results, protein_viz_data){
  
  # get items 
  df <- enr_results
  df$ITEMS <- (lapply(df$ITEMS, fix_string_list))
  inverse_enr = as.data.frame(df) %>% unnest(c("ITEMS")) 
  
  inverse_enr = inverse_enr %>% group_by(ITEMS) %>%
    summarize(GENE.SETS = list(GENE.SET),
              average.NES = mean(NES),
              average.ENR_PVAL = mean(PVAL),
              average.ENR_ADJ.PVAL = mean(ADJ.PVAL))
  
  inverse_enr <- inverse_enr[order(inverse_enr$average.ENR_PVAL),]
  
  # return only protein inside the experiment comparison
  inverse_enr[inverse_enr$ITEMS %in% protein_viz_data$GeneName,]
}

df = results2$`Kidney - Muscle`$Biological_process
head(inverse_enr)

inverse_enr[inverse_enr$ITEMS %in% protein_viz$data[[1]]$GeneName,]
```



# Add acceptance test

```{r}

library(EnrichmentAnalysisStepR)
experiment_home = "../test_data/"
results = enrichment_workflow_step(experiment_home, 
                                      gmt_folder = experiment_home,  
                                      output_folder = experiment_home,
                                      by = "Gene")

expected_data = read_json("../test_data/C - P/expected_Pathway.json")
```


# Dev Writer:

```{r}
getwd()
```


```{r}


enrich_write_output_tables <- function(results, output_folder){
  for (comparison in names(results)){
    first_condition = strsplit(comparison,"\\s+")[[1]][1]
    second_condition = strsplit(comparison,"\\s+")[[1]][3]
    for (library in names(results[[comparison]])){
      file_name = str_c(str_c(first_condition, second_condition,sep = "."),
                        library,sep = ".")
      #print(file_name)
      enr = list(up.condition = unbox(first_condition),
                 down.condition = unbox(second_condition),
                 database = unbox(library),
                 version = unbox("0.0.0"),
                 gene.set.results = as.data.frame(results[[comparison]][[library]]))
      
      print(str_c(file_name,".json"))
      write_json(enr,file.path(output_folder,str_c(file_name,".json")))
    }
  }
}

string_to_array <- function(bad_string){
  bad_string <- gsub("'","",bad_string)
  bad_string <- gsub("\\[","",bad_string)
  bad_string <- gsub("\\]","",bad_string)
  vec <- str_split(bad_string, ", ")
  vec
}

enrich_write_output_tables(results, "./test")


```


```{r}
test_results <- results
test_results$`C - P`$BiologicalProcess$items <- lapply(unlist(test_results$`C - P`$BiologicalProcess$items), string_to_array)
head(test_results$`C - P`$BiologicalProcess)
```

```{r}





```


proof of concept for using gs to parse genes in gene sets

```{r}

add_gene_sets_to_result <- function(result_table, gs){
  result_table$items <- gs[match(result_table$gene.set, names(gs))]
  result_table
}

gmt.file = "../../../experiments/BKG_gmt_output_human_demo/9606/BiologicalProcess.gmt"
gs <- getGenesets(gmt.file)
  
results$`C - P`$BiologicalProcess <- add_gene_sets_to_result(results$`C - P`$BiologicalProcess, gs)

#results$`C - P`$BiologicalProcess$items
```
```{r}

enrich_write_output_tables <- function(results, output_folder){
  for (comparison in names(results)){
    first_condition = strsplit(comparison,"\\s+")[[1]][1]
    second_condition = strsplit(comparison,"\\s+")[[1]][3]
    for (library in names(results[[comparison]])){
      file_name = str_c(str_c(first_condition, second_condition,sep = "."),
                        library,sep = ".")
      #print(file_name)
      enr = list(up.condition = unbox(first_condition),
                 down.condition = unbox(second_condition),
                 database = unbox(library),
                 version = unbox("0.0.0"),
                 gene.set.results = as.data.frame(results[[comparison]][[library]]))
      
      print(str_c(file_name,".json"))
      write_json(enr,file.path(output_folder,str_c(file_name,".json")))
    }
  }
}

enrich_write_output_tables(results, "./test")

```


# Debug Workflow error Her:

error message " "


```{r small experiment}
library(EnrichmentAnalysisStepR)
experiment_home = "/home/joseph/experiments/her_enrichment_bug_test/"
gmt_folder = "/home/joseph/experiments/BKG_gmt_output_human/9606"
results = enrichment_workflow_step(experiment_home, 
                                   gmt_folder = gmt_folder,
                                   output_folder = file.path(experiment_home,"transform"),
                                   by = "Protein")
```

```{r}
upload_folder = experiment_home
by = "Protein"
method = "gsea"
# Get Experiment Data
protein_viz_path = file.path(upload_folder,"protein_viz.json")
stopifnot(file.exists(protein_viz_path))
protein_viz = read_json(protein_viz_path, simplifyVector = TRUE)
stopifnot(dim(protein_viz)[1]>0)

protein_count_and_intensities = read_json(file.path(upload_folder,"protein_counts_and_intensity.json"), simplifyVector = TRUE)
protein_count_and_intensities = process_protein_intensities(protein_count_and_intensities, by = by)
stopifnot(dim(protein_count_and_intensities)[1]>0)

results = list()
for (row in (1:nrow(protein_viz))){

  comparison = protein_viz[row,"conditionComparison"]
  print(comparison)

  comparison_viz = protein_viz[row, "data"]
  tmp_protein_int = filter_protein_ints(comparison, protein_count_and_intensities)

  cls_vec = get_cls_vec(comparison, tmp_protein_int)
  se_comparison <- protein_viz_int_2_de_exp(comparison_viz, tmp_protein_int, cls_vec, by = by)

  results[[comparison]] = perform_comparison_enrichment(se_comparison, gmt_folder, method)

}
```


# Try to get signal to noise ratio and or ranks

```{r}
library(EnrichmentAnalysisStepR)

upload_folder= "/home/joseph/experiments/her_data_gsea_test"
gmt_folder = "/home/joseph/experiments/BKG_gmt_output_human/9606"
by = "Protein"

# Get Experiment Data
protein_viz_path = file.path(upload_folder,"protein_viz.json")
stopifnot(file.exists(protein_viz_path))
protein_viz = read_json(protein_viz_path, simplifyVector = TRUE)
stopifnot(dim(protein_viz)[1]>0)

protein_count_and_intensities = read_json(file.path(upload_folder,"protein_counts_and_intensity.json"), simplifyVector = TRUE)
protein_count_and_intensities = process_protein_intensities(protein_count_and_intensities, by = by)
stopifnot(dim(protein_count_and_intensities)[1]>0)

# for comparison in comparisons...

results = list()
for (row in (1:nrow(protein_viz))){

  comparison = protein_viz[row,"conditionComparison"]
  print(comparison)

  comparison_viz = protein_viz[row, "data"]
  tmp_protein_int = filter_protein_ints(comparison, protein_count_and_intensities)

  cls_vec = get_cls_vec(comparison, tmp_protein_int)
  se_comparison <- protein_viz_int_2_de_exp(comparison_viz, tmp_protein_int, cls_vec, by = by)

  #results[[comparison]] = perform_comparison_enrichment(se_comparison, gmt_folder, method)

}

## compile output tables by merging enrichment results with annotation data
#results <- enrichment_adjust_p_values(results)
#results <- enrichment_annotate_results(results, gmt_folder)
#enrich_write_output_tables(results, output_folder)

```

```{r}
comparison = se_comparison
enrichment = list()

for (library in list.files(gmt_folder, pattern = "\\.gmt$")){
  gmt.file <- file.path(gmt_folder, library)

  gs <- getGenesets(gmt.file)

  print(gmt.file)
  print(paste("Sets in Library: ", length(gs)))
  print(paste("Genes in Library: ", length(unique(unlist(gs)))))
  print(paste("Genes in Library and dataset: ", length(intersect(rownames(comparison),unlist(gs)))))

  # check for common ids
  if (length(intersect(rownames(comparison),unlist(gs)))>100){

    # do enrichment
    #sbea.res <- sbea(method = method, se = comparison, 
    #                 gs = gmt.file, perm = 250, alpha = 1)
    #results = gsRanking(sbea.res)

    # get sets info (size/observed and merge)
    #sets_info <- add_sets_info(gs, comparison)
    #results <- merge(results,sets_info)

    # write results
    #lib = str_split(library,".gmt")[[1]][1]
    #label = str_c(lib,"_enrichment.csv")

    #enrichment[[lib]] = add_gene_sets_to_result(results, gs)
    
  } else {
    print("Skipping enrichment, not enough genes in common.")
    print("Heres some example ids")
    print(rownames(comparison)[1:10])
    print(unlist(gs)[1:10])
  }

}

enrichment
```
```{r}
se = se_comparison
gs = gs
perm = 1000

padog <- NULL
isAvailable("PADOG", type="software")

grp <- se[[configEBrowser("GRP.COL")]]
grp <- ifelse(grp == 0, "c", "d") 

blk <- NULL
BLK.COL <- configEBrowser("BLK.COL")
if(BLK.COL %in% colnames(colData(se))) 
    blk <- make.names(colData(se)[,BLK.COL]) 
paired <- !is.null(blk)

nmin <- configEBrowser("GS.MIN.SIZE")
perm <- as.numeric(perm) 

res <- padog(assay(se), group=grp, 
    paired=paired, block=blk, gslist=gs, Nmin=nmin, NI=perm)
```

# Try to Get intermediate GSEA stats

```{r}
source("./external_code_do_not_commit/GSEA1.R")

se = se_comparison
GRP.COL = "GROUP"
call.nperm = 250

# get inputs to gene ranking
A = as.data.frame(assay(se_comparison))
head(A)

cls <- list()
cls$phen <- levels(as.factor(se[[GRP.COL]]))
cls$class.v <- ifelse(se[[GRP.COL]] == cls$phen[1], 0, 1)
class.labels = cls$class.v 
print(class.labels)

gene.labels = rownames(se_comparison)
#print(gene.labels)

O <- GSEA.GeneRanking(as.matrix(A), class.labels, gene.labels, call.nperm)
#order.matrix[,n.starts[nk]:n.ends[nk]] <- O$order.matrix
#obs.order.matrix[,n.starts[nk]:n.ends[nk]] <- O$obs.order.matrix
#correl.matrix[,n.starts[nk]:n.ends[nk]] <- O$s2n.matrix
#obs.correl.matrix[,n.starts[nk]:n.ends[nk]] <- O$obs.s2n.matrix
#rm(O)
apply(O$obs.s2n.matrix, 1, median)
#O$s2n.matrix


obs.s2n <- apply(obs.correl.matrix, 1, median)    
names(obs.s2n) <- gene.labels
save(obs.s2n, file=file.path(output.directory, "gsea_s2n.RData"))  
      
obs.index <- order(obs.s2n, decreasing=TRUE)            
obs.s2n <- obs.s2n[obs.index]       
```


```{r}
# inputs

se = se_comparison
gs.db = gs
nperm = 250
alpha = 1


Ng <- length(gs.db)
gs.names <- names(gs.db)
size.G <- sapply(gs.db, length) 
gs <- matrix(NA, nrow=Ng, ncol=max(size.G))
for(i in seq_len(Ng)) gs[i, seq_len(size.G[i])] <- gs.db[[i]]

N <- nrow(A)
Ns <- ncol(A)
#all.gene.descs <- all.gene.symbols <-  gene.labels[i]

Obs.indicator <- Obs.RES <- matrix(nrow= Ng, ncol=N)
Obs.ES <- Obs.arg.ES <- Obs.ES.norm <- vector(length = Ng, mode = "numeric")

# Compute observed and random permutation gene rankings
obs.s2n <- vector(length=N, mode="numeric")
signal.strength <- tag.frac <- gene.frac <- 
  coherence.ratio <- vector(length=Ng, mode="numeric")
obs.phi.norm <- matrix(nrow = Ng, ncol = nperm)
correl.matrix <- obs.correl.matrix <- order.matrix <- 
  obs.order.matrix <- matrix(nrow = N, ncol = nperm)

nperm.per.call <- 100
n.groups <- nperm %/% nperm.per.call
n.rem <- nperm %% nperm.per.call
n.perms <- c(rep(nperm.per.call, n.groups), n.rem)
n.ends <- cumsum(n.perms)
n.starts <- n.ends - n.perms + 1
n.tot <- ifelse(n.rem == 0, n.groups, n.groups + 1)

for (nk in seq_len(n.tot)) 
{
  call.nperm <- n.perms[nk]
  message(paste("Permutations:", n.starts[nk], "--", n.ends[nk]))
  O <- GSEA.GeneRanking(as.matrix(A), class.labels, gene.labels, call.nperm)
  order.matrix[,n.starts[nk]:n.ends[nk]] <- O$order.matrix
  obs.order.matrix[,n.starts[nk]:n.ends[nk]] <- O$obs.order.matrix
  correl.matrix[,n.starts[nk]:n.ends[nk]] <- O$s2n.matrix
  obs.correl.matrix[,n.starts[nk]:n.ends[nk]] <- O$obs.s2n.matrix
  rm(O)
}

message("Processing ...")
# using median to assign enrichment scores
obs.s2n <- apply(obs.correl.matrix, 1, median)    
names(obs.s2n) <- gene.labels
save(obs.s2n, file=file.path(output.directory, "gsea_s2n.RData"))  

obs.index <- order(obs.s2n, decreasing=TRUE)            
obs.s2n <- obs.s2n[obs.index]       
```


```{r}
rnks = as.data.frame(cbind(gene.labels, obs.index))
colnames(rnks) = c("genes", "ranks")
write.table(rnks, file = "ranks.rnk", sep = "\t", quote = FALSE, row.names = F)
```

```{r}

intersect(unlist(gs)[1:1000], gene.labels)
```


```{r}
write_gsea_expression_file <- function(se_comparison){
  expression <- as.data.frame(assay(se_comparison))
  expression <- tibble::rownames_to_column(expression, "NAME")
  expression <- tibble::add_column(expression, "DESCRIPTION" = "NA", .after = "NAME")
  write.table(expression, file = "gsea_expression_file.txt", sep = "\t", quote = FALSE, row.names = F)
}

write_gsea_expression_file(se_comparison)
```

```{r}
se_comparison$Treatment
n_samples = length(se_comparison$Treatment)
n_conditions = length(unique(se_comparison$Treatment))
paste(unique(se_comparison$Treatment))
se_comparison$Treatment


```
```{r}


```


```{r}
write_gsea_cls_file <- function(se_comparison){
  sink("gsea_cls_file.cls")
  
  #first line
  cat(length(se_comparison$Treatment))
  cat(" ")
  cat(length(unique(se_comparison$Treatment)))
  cat(" 1\n")
  
  #second line
  cat("# ")
  for (condition in unique(se_comparison$Treatment)){
    cat(condition)
    cat(" ")
  }
  cat("\n")
  
  #third line
  for (condition in (se_comparison$Treatment)){
    cat(condition)
    cat(" ")
  }
  cat("\n")
  sink()
}

write_gsea_cls_file(se_comparison)

```


