---
title: "Mass Dynamics enrichment service"
output:
  github_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r eval=FALSE}
devtools::install_github("MassDynamics/EnrichmentAnalysisStepR")
```

```{r}
library(EnrichmentAnalysisStepR)
utils::packageVersion("EnrichmentAnalysisStepR")
```

# Workflow example

The example runs the differential expression analysis using the [HER2 dataset](https://app.massdynamics.com/experiments/papers/c6fc6c60-fe65-47cb-bd6d-a021f0ed8720#/volcano) (https://pubmed.ncbi.nlm.nih.gov/26883193/) whose data is included in the Mass Dynamics [MassExpression](https://github.com/MassDynamics/MassExpression) R package. The differential expression results are then used as input for the enrichment step.  

## Differential expression analysis with `MassExpression`

```{r include=FALSE}
devtools::install_github("MassDynamics/MassExpression")
```

```{r}
library(MassExpression)
```


```{r}
intensities <- mq_lfq_data$intensities
design <- mq_lfq_data$design
parameters <- mq_lfq_data$parameters
normalisation_method <- parameters[parameters[,1] == "UseNormalisationMethod",2]
species <- parameters[parameters[,1] == "Species",2]
labellingMethod <- parameters[parameters[,1] == "LabellingMethod",2]


results <- runGenericDiscovery(experimentDesign = design, 
                               proteinIntensities = intensities, 
                               normalisationMethod = normalisation_method, 
                               species = species, 
                               labellingMethod = labellingMethod)

# Save results
IntensityExperiment <- results$IntensityExperiment
CompleteIntensityExperiment <-  results$CompleteIntensityExperiment
longIntensityDT <- results$longIntensityDT

saveOutput(IntensityExperiment = IntensityExperiment, 
CompleteIntensityExperiment = CompleteIntensityExperiment,
longIntensityDT = longIntensityDT, 
outputFolder =  "example_data/testDE")
```

## Run enrichment with `EnrichmentAnalysisStepR`

As an example, the GMT file for Human Gene Ontologies, used in in the Mass Dynamics enrichment service, is added to this repository in folder `example_data/9606-Human`.  

```{r}
gmt_folder <- "example_data/9606-Human"
results = runEnrichmentWorkflowStep(uploadFolder = "example_data/testDE",
                                   gmtFolder = gmt_folder,
                                   outputFolder = "example_data/testEnrichment",
                                   by = "Protein",
                                   method = "camera")
```


## Access enrichment results

`results` is a list of lists containing the results. Each entry of the `results` list contains the enrichment results for one pairwise comparison.

- Pairwise comparisons tested 

```{r}
names(results)
```

- Gene sets tested

```{r}
names(results[[1]]$libraryStatistics)
```

- Enrichment results for **GO Biological Processes**

```{r}
head(results[[1]]$libraryStatistics$BiologicalProcess)[,c("gene.set", "afc","pval", "adj.pval", "observed", "name")]
```

## Plot enrichment volcano plot

```{r}
library(ggplot2)
library(ggrepel)
```

This results are the same obtained in the [HER2 demo experiment](https://app.massdynamics.com/experiments/papers/c6fc6c60-fe65-47cb-bd6d-a021f0ed8720#/enrichment-analysis) in the Mass Dynamics platform.  

```{r}
results_df <- results[[1]]$libraryStatistics$BiologicalProcess
ggplot(results_df, aes(x = afc, y = -log10(pval))) + geom_point() + theme_bw() +
  geom_text_repel(data = results_df[-log10(results_df$pval) > 4.7, ], aes(x = afc, y = -log10(pval), label = name)) + 
  ggtitle("Biological Processes")
```

